Extended Data

### 2021-10-19 doc/manuscript/Dow_manuscript_Nature.Rmd
First go at author contributions

 <!--Climate -->
Climate data corresponding to the measurement periods at both sites were obtained from local weather stations. 
For SCBI, weather data were obtained from a meteorological tower adjacent to the SCBI ForestGEO plot, via the ForestGEO Climate Data Portal v1.0 (https://forestgeo.github.io/Climate/)[@anderson-teixeira_forestgeo_2020]. 
Daily temperature records for Harvard Forest were obtained from the Harvard Forest weather station **(Boose & Gould, 2004; Boose, 2018)** (**CAM, we need these references. And see [GitHub issue #81](https://github.com/EcoClimLab/growth_phenology/issues/81)**).
The R package *clim.pact*[@climpact_REF; @benestad_empiricalstatistical_2008] was used to plot temperatures for visual inspection and to identify readings that were >3 standard deviations away from yearly means, which were labeled as outliers and removed from the dataset. 
Gaps in the SCBI meteorological tower data were subsequently filled using temperature readings obtained from a National Center for Environmental Information (NCEI) weather station located in Front Royal, Virginia (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USC00443229/detail). 
For each site, we extracted daily maximum ($T_{max}$) and minimum temperatures ($T_{min}$).

@@

<!--QUILT PLOT METHODS-->
Correlations between monthly climate and $RWI$ were assessed using ‘dplR’[@bunn_dendrochronology_2008] and ‘bootRes’[@zang_dendroclimatic_2013] in R v.**##** (R Core Team, **YEAR**), which correlated functions and bootstrapped confidence intervals for these relationships[@biondi_dendroclim2002_2004].
We analyzed these correlations for January through September of the current year (presented in Fig. `r tree_rings`, `r EDF_tree_rings_TMIN`).
To test for potential lag effects effects of spring temperatures on the following year's growth, we also ran a version of the analysis extending back to include climate of every month of the previous year (`r EDF_tree_rings_TMAX_extended`).
Correlations and significance levels for months April-August are given in **APPENDIX**.

<!--MULTIVARIATE ANALYSIS METHODS-->
We used a multivariate model to test for joint effects of April and summer $T_{max}$ on RWI.
We began by testing univariate correlations of $T_{max}$ over three summer windows: June, June-July, and May-August. 
Having determined that, among these, June-July explained the most variation, we then analyzed the joint effects of April $T_{max}$ and June-July $T_{max}$ on RWI for each chronology independently using **(Function)** in R. 
Slopes and p-values for each chronology are given in **APPENDIX**.

@@

**(CAM, please draft this before sending around.)**

>>>>

 <!--Climate -->
Climate data corresponding to the measurement periods at both sites were obtained from local weather stations. 
For SCBI, weather data were obtained from a meteorological tower adjacent to the SCBI ForestGEO plot, via the ForestGEO Climate Data Portal v1.0 (https://forestgeo.github.io/Climate/)[@anderson-teixeira_forestgeo_2020]. 
Daily temperature records for Harvard Forest were obtained from the Harvard Forest weather station[@boose_shaler_2021-1; @boose_fisher_2021-2]. Climate data for Harvard Forest were complete and used as is. 
The R package *clim.pact*[@climpact_REF; @benestad_empiricalstatistical_2008] was used to plot temperatures for visual inspection and to identify readings that were >3 standard deviations away from yearly means, which were labeled as outliers and removed from the dataset. 
Gaps in the SCBI meteorological tower data were subsequently filled using temperature readings obtained from a National Center for Environmental Information (NCEI) weather station located in Front Royal, Virginia (https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USC00443229/detail). 
For each site, we extracted daily maximum ($T_{max}$) and minimum temperatures ($T_{min}$).

@@

<!--QUILT PLOT METHODS-->
Correlations between monthly climate and $RWI$ were assessed using ‘dplR’[@bunn_dendrochronology_2008] and ‘bootRes’[@zang_dendroclimatic_2013] in R v.1.3.1093 (R Core Team, 2020), which correlated functions and bootstrapped confidence intervals for these relationships[@biondi_dendroclim2002_2004].
We analyzed these correlations for January through September of the current year (presented in Fig. `r tree_rings`, `r EDF_tree_rings_TMIN`).
To test for potential lag effects effects of spring temperatures on the following year's growth, we also ran a version of the analysis extending back to include climate of every month of the previous year (`r EDF_tree_rings_TMAX_extended`).
Correlations and significance levels for months April-August are given in **APPENDIX**.

<!--MULTIVARIATE ANALYSIS METHODS-->
We used a multivariate model to test for joint effects of April and summer $T_{max}$ on RWI.
We began by testing univariate correlations of $T_{max}$ over three summer windows: June, June-July, and May-August. 
Having determined that, among these, June-July explained the most variation, we then analyzed the joint effects of April $T_{max}$ and June-July $T_{max}$ on RWI for each chronology independently using the base lm() function in R. 
Slopes and p-values for each chronology are given in **APPENDIX**.

@@

All authors gave suggestions on improvements to manuscript during development.

Albert Y. Kim - Developed and assisted in coding the statistical framework for analyses used.

Loïc D’Orangeville - Provided tree core chronologies for additional sites, gave feedback on draft of manuscript.

Erika B. Gonzalez-Akre - Oversees and manages data collection at SCBI.

Ryan Helcoski - Collected and measured cores at SCBI.

Valentine Herrmann - Assisted with coding and interpretation of statistical results.

Grant L. Harley - Provided tree core chronologies for additional sites and assisted with interpretation of results.

Justin T. Maxwell - Provided tree core chronologies for additional sites, detrended crns, and assisted with interpretation of results.

Ian R. McGregor - Assisted with the acquisition and processing of leaf phenology data.

William McShea -

Sean McMahon -

David A. Orwig -

Neil Pederson - Provided tree core chronologies for additional sites and gave feedback on tree core analysis

Alan J. Tepley -

Kristina J. Anderson-Teixeira - Oversaw all aspects of research and manuscript development.
